---
title: "README"
author: "Steffi LaZerte"
output: github_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(weathercan)
library(dplyr)
```

# weathercan

[![Build Status](https://travis-ci.org/steffilazerte/weathercan.svg?branch=master)](https://travis-ci.org/steffilazerte/weathercan) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/steffilazerte/weathercan?branch=master&svg=true)](https://ci.appveyor.com/project/steffilazerte/weathercan)

This package is makes it easier to search for and download multiple months/years of historical weather data from Environment and Climate Change Canada (ECCC) website.

Bear in mind that these downloads can be fairly large and performing multiple downloads may use up ECCC's bandwidth unecessarily. Try to stick to what you need.

For more details and tutorials checkout the [weathercan website](http://steffilazerte.github.io/weathercan)

## Installation

Use the `devtools` package to directly install R packages from github:

```{r, eval = FALSE}
install.packages("devtools") # If not already installed
devtools::install_github("steffilazerte/weathercan") 
```

## General usage

To download data, you first need to know the `station_id` associated with the station you're interested in.

### Stations

`weathercan` includes a data frame called `stations` which includes a list of stations and their details (including `station_id`.

```{r}
head(stations)
glimpse(stations)
```

You can look through this data frame directly, or you can use the `stations_search` function:

```{r}
stations_search("Kamloops", interval = "hour")
```

Time frame must be one of "hour", "day", or "month".

You can also search by proximity:

```{r}
stations_search(coords = c(50.667492, -120.329049), dist = 20, interval = "hour")
```

### Weather

Once you have your `station_id`(s) you can download weather data:

```{r}
kam <- weather(station_ids = 51423, start = "2016-01-01", end = "2016-02-15")
kam
```

You can also download data from multiple stations at once:

```{r}
kam_pg <- weather(station_ids = c(48248, 51423), start = "2016-01-01", end = "2016-02-15")
kam_pg
```

And plot it:

```{r, fig.height=6, fig.width=12}
library(ggplot2)

ggplot(data = kam_pg, aes(x = time, y = temp, group = station_name, colour = station_name)) +
  theme(legend.position = "top") +
  geom_line() +
  theme_minimal()
```
