---
title: "README"
author: "Steffi LaZerte"
date: "June 21, 2016"
output: md_document
---

```{r, echo = FALSE}
library(envirocan)
```

# envirocan

This package is makes it easier to search for and download multiple months/years of historical weather data from Environment Canada webiste.

## Installation

Use the `devtools` package to directly install R packages from github:

```{r, eval = FALSE}
install.packages("devtools") # If not already installed
devtools::install_github("steffilazerte/envirocan", ref = "v0.1.1.1") 
## For most recent release; Otherwise omit "ref = " to download most recent version
```

## Basic usage:

To download data, you first need to know the `station_id` associated with the station you're interested in.

### Stations

`envirocan` includes a data frame called `stations` which includes a list of stations and their details (including `station_id`.

```{r}
head(stations)
```

You can look through this data frame directly, or you can use the `stations_search` function:

```{r}
stations_search("Kamloops", timeframe = "hour")
```

Time frame must be one of "hour", "day", or "month".

You can also search by proximity:

```{r}
stations_search(coords = c(50.667492, -120.329049), dist = 20, timeframe = "hour")
```

### Weather

Once you have your `station_id`(s) you can download weather data:

```{r}
kamloops <- weather(station_ids = 51423,
                    start = "2016-01-01", end = "2016-02-15")
                    
head(kamloops)

```

You can also download data from multiple stations at once:

```{r}
kam.pg <- weather(station_ids = c(48248, 51423),
                  start = "2016-01-01", end = "2016-02-15")
                    
head(kam.pg)
```

And plot it:

```{r}
library(ggplot2)

ggplot(data = kam.pg, aes(x = time, y = temp, group = station_name, colour = station_name)) +
  theme(legend.position = "top") +
  geom_line()
```
