---
title: "The meanings of flags"
author: "Steffi LaZerte"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flags}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, message = FALSE}
library(weathercan)
library(dplyr)
```


## What are flags

The data output of the `weather_dl()` function include a corresponding `_flag` column for each data column. These columns are used by ECCC to add notes regarding measurements. 

In the `weather_dl()` function if `format = TRUE` (the default), data corresponding to flags `M`, `NA`, `[empty]` and `L` are all replaced with `NA`.

For example, a sample of unformatted data from Magog station in Quebec looks like:

```{r, echo = FALSE}
options(width = 100)
w <- weather_dl(station_id = 5401, start =  "2017-01-01", interval = "month", format = FALSE)

dplyr::select(w, station_name, `Date/Time`, `Total Precip (mm)`, `Total Precip Flag`) %>% tail()
dplyr::select(w, station_name, `Date/Time`, `Snow Grnd Last Day (cm)`, `Snow Grnd Last Day Flag`) %>% tail()
```

In this output, you can see two flags: `^` in `Total Precip` and `M` in `Snow Grnd Last Day`

This same sample, formatted looks like:

```{r, echo = FALSE}

weather_dl(station_id = 5401, start =  "2017-01-01", interval = "month") %>%
  dplyr::select(date, total_precip, total_precip_flag, snow_grnd_last_day, snow_grnd_last_day_flag) %>%
  tail()
```

As you can see, we still have the two flags, but the missing data flag (`M`) is now replaced with NA. The other flag `^` is not, as it indicates that "The value displayed is based on incomplete data" (see below).

## Flags

The flags index can be accessed through the built in data frame: `flags`

```{r, echo = FALSE}
knitr::kable(flags)
```



