---
title: "The meanings of flags"
author: "Steffi LaZerte"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flags}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message = FALSE}
library(weathercan)
library(dplyr)
```


## What are flags

The data output of the `weather()` function include a corresponding `_flag` column for each data column. These columns are used by ECCC to add notes regarding measurements. 

In the `weather()` function if `format = TRUE` (the default), data corresponding to flags `M`, `NA`, `[empty]` and `L` are all replaced with `NA`.

For example, a sample of unformated data from Magog station in Quebec looks like:

```{r, echo = FALSE}
weather(station_id = 5401, start =  "2017-01-01", interval = "month", format = FALSE) %>%
  select(station_name, station_id, `Date/Time`, `Total Precip (mm)`, `Total Precip Flag`, `Snow Grnd Last Day (cm)`, `Snow Grnd Last Day Flag`) %>% 
  tail()
```

This same sample, formated looks like:

```{r, echo = FALSE}
weather(station_id = 5401, start =  "2017-01-01", interval = "month") %>%
  select(station_name, station_id, date, total_precip, total_precip_flag, snow_grnd_last_day, snow_grnd_last_day_flag) %>%
  tail()
```

As you can see, the missing data flag (`M`) is now replaced with NA. The other flag `^` is not, as it indicates that "The value displayed is based on incomplete data" (see below).

## Flags

```{r}
knitr::kable(flags)
```



