---
title: "Climate Normals"
author: "Steffi LaZerte"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Climate Normals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
old <- options(tibble.max_extra_cols = 0, width = 90)
```

## Downloading Climate Normals

Climate Normals and Averages describe the average climate conditions specific to a particular location. These can be downloaded from Environment and Climate Change Canada using the `normals_dl()` function.

First we'll load the `weathercan` package for downloading the data and the `tidyr` package for unnesting the data (see below).

```{r setup, message = FALSE}
library(weathercan)
library(tidyr)
library(dplyr)
library(naniar) # For exploring missing values
```

To download climate normals, we'll first find the stations we're interested in using the `stations_search()` function. We'll use the `normals_years = "current"` argument to filter to only stations with available climate normals for the `1981-2010` year range.

```{r}
stations_search("Winnipeg", normals_years = "current")
```

Let's look at the climate normals from this station in Winnipeg, MB. Note that unlike the `weather_dl()` function, the `normals_dl()` function requires `climate_id` not `station_id`. By default the normals are downloaded for the years "1981-2010" (currently 1981-2010 and 1971-2000 are the only year ranges available)

```{r}
n <- normals_dl(climate_ids = "5023222")
n
```

Because there are two different types of climate normals (weather measurements and first/last frost dates), the data are nested as two different datasets. We can see that the Airport (Richardson Int'l) has `r ncol(n$normals[[1]])` average weather measurements/codes as well as first/last frost dates.

We can also see that this station has data quality sufficient to meet the WMO standards for temperature and precipitation (i.e. both these measurements have code >= A). See the [ECCC calculations document](https://climate.weather.gc.ca/doc/Canadian_Climate_Normals_1981_2010_Calculation_Information.pdf) for more details.

To extract either data set we can use the `unnest()` function from the `tidyr` package.

```{r}
normals <- unnest(n, normals)
frost <- unnest(n, frost)
```

Note that this extracts the measurements for all three stations (in the case of the `normals` data frame), but not all measurements are available for each station

```{r, R.options = list(tibble.print_max = Inf)}
normals
```

To visualize missing data we can use the `gg_miss_var()` function from the `naniar` package.
```{r, eval = FALSE, fig.asp = 2, out.width = "100%", fig.width = 8}
select(normals, -contains("_code")) %>%  # Remove '_code' columns
  gg_miss_var(facet = station_name)
```

```{r, warning = FALSE, fig.asp = 2, out.width = "100%", fig.width = 8}
suppressWarnings({select(normals, -contains("_code")) %>%  # Remove '_code' columns
    gg_miss_var(facet = station_name)})
```

Let's take a look at the frost data.

```{r, R.options = list(width = 90)}
if("normals" %in% names(frost)) frost <- select(frost, -normals) # tidyr v1
glimpse(frost)
```

### Finding stations with specific measurements

The include data frame, `normals_measurements` contains a list of stations with their corresponding measurements. Be aware that this data might be out of date!

```{r}
normals_measurements
```

For example, if you wanted all `climate_id`s for stations that have data on 
soil temperature for 1981-2010 normals:

```{r}
normals_measurements %>%
  filter(stringr::str_detect(measurement, "soil"),
         normals == "1981-2010") %>%
  pull(climate_id) %>%
  unique()
```

## Understanding Climate Normals

The measurements contained in the climate normals are very specific. To better understand how they are calculated please explore the following resources:

- ECCC Climate Normals Calculations ([1981-2010](https://climate.weather.gc.ca/doc/Canadian_Climate_Normals_1981_2010_Calculation_Information.pdf) | [1971-2000](https://climate.weather.gc.ca/doc/Canadian_Climate_Normals_1971_2000_Calculation_Information.pdf))
    - [`weathercan` Climate Normals Codes](flags.html)
- [ECCC Climate Normals Technical Documentation](https://www.canada.ca/en/environment-climate-change/services/climate-change/canadian-centre-climate-services/display-download/technical-documentation-climate-normals.html)
    - [`weathercan` Climate Normals Glossary](glossary_normals.html)


```{r, include = FALSE}
# Reset options
options(old)
```
