---
title: "Meteoland"
author: "Steffi LaZerte"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
---

[`meteoland`](https://github.com/vegmod/meteoland) is a package for interpolating meteorological data over spatial scales. As of version v0.7.9 they support transforming and important `weathercan` output for use in spatial interpolations. In this article we will go over a hypothetical example.

## Setup
First we'll load the packages and find the stations we're interested in and 
```{r}
library(meteoland)
library(weathercan)
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(rnaturalearth)

s <- stations %>%
  filter(prov == "MB", interval != "month", start <= 2015, end >= 2015, !is.na(elev)) %>%
  group_by(station_id) %>%
  mutate(n = length(interval)) %>%
  filter(n == 2)

s_map <- s %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(3347)

mb <- ne_states(country = "Canada", returnclass = "sf") %>%
  filter(name == "Manitoba") %>%
  st_transform(3347)

ggplot() +
  geom_sf(data = mb) +
  geom_sf(data = s_map)
```

Let's focus on northern Manitoba

```{r}
s <- filter(s, lat > 55)
s_map <- s %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
  st_transform(3347)

ggplot() +
  geom_sf(data = mb) +
  geom_sf(data = s_map)
```


## Download data

```{r}
mb_hr <- weather_dl(station_id = unique(s$station_id), interval = "hour",
                    start = "2015-01-01", end = "2015-12-31", verbose = TRUE)
mb_day <- weather_dl(station_id = unique(s$station_id), interval = "day",
                     start = "2015-01-01", end = "2015-12-31", verbose = TRUE)
```

## Pass to `meteoland`

First we'll reshape our hourly and daily data into a `meteoland` interpolations data object.
```{r}
mb_north <- reshapeweathercan(mb_hr, mb_day, output = "MeteorologyInterpolationData")
```

We can get a sense of the data coverage (number of dates with data per station per variable).
```{r}
interpolation.coverage(mb_north, type = 'spatial') %>%
  head()
```

Or the number of stations with data per date per variable.
```{r}
interpolation.coverage(mb_north, type = 'temporal') %>%
  head()
```

Next we have to calibrate the variable we're intersted in (here `Tmin`)
```{r}
tmin <- interpolation.calibration(mb_north, variable = "Tmin",
                                  N_seq = 20,
                                  alpha_seq = seq(5, 10, by = 1),
                                  verbose = TRUE)

mb_north@params$N_MinTemperature = tmin$N
mb_north@params$alpha_MinTemperature = tmin$alpha
```

Next we cross-validate the data
```{r}
cv <- interpolation.cv(mb_north, verbose = TRUE)
summary(cv)
```

We create a dummy `SpatialPointsTopography` object representing the points (in this case) that we wish to interpolate over. Note that here I'm using a mean elevation as a placeholder as I don't have actual elevation values. If you have slope and aspect, even better. Remember, interpolation is only as good as the data you give it!

```{r}
interp <- expand.grid(lat = seq(56, 58.5, 0.25),
                      lon = seq(-101, -95, 0.25), 
                      elev = mean(s_map$elev))

interp <- SpatialPointsTopography(as.matrix(interp[, c("lon", "lat")]),
                                  elevation = interp$elev,
                                  proj4string = CRS("+proj=longlat +ellps=WGS84"))
```

Now for the actual interpolation. Here we interpolate over all dates in the range of the original data.
```{r}
new_interp <- interpolationpoints(mb_north, interp)
```

We can summarize and plot our interpolations
```{r}
map <- summarypoints(new_interp, var = "MinTemperature")

mb_interp <- st_as_sf(map) %>%
  rename_at(.vars = vars(contains("matrix")), ~"min_temp") %>%
  mutate(station_id = NA, interp = TRUE)

mb_stations <- mb_day %>%
  group_by(station_id, lat, lon) %>%
  summarize(min_temp = mean(min_temp, na.rm = TRUE)) %>%
  st_as_sf(coords = c("lon", "lat"), crs = st_crs(mb_interp)) %>%
  mutate(interp = FALSE)

map_sf <- rbind(mb_interp, mb_stations) %>%
  mutate(interp = factor(interp, levels = c("TRUE", "FALSE")))

ggplot() +
  geom_sf(data = mb) +
  geom_sf(data = map_sf, aes(fill = min_temp, shape = interp, colour = interp), size = 4) +
  scale_shape_manual(values = c(21,23)) +
  coord_sf(ylim = c(2100000, 2550000))
```
